### 2016-02-08 text/Chapter04.tex
Chapter 5 intro and semantics

The development in this chapter can be seen as a counterpart to the well-known development of 
\emph{effect systems} \cite{effects-gifford}. The Chapter~\ref{ch:transl} then links \emph{coeffects} 
with \emph{comonads} in the same way in which effect systems can be linked with monads
\cite{monad-notions}. The syntax and type system of the flat coeffect calculus follows 
a similar style as effect systems \cite{effects-polymorphic,effects-talpin-et-al}, but differs
in the structure, as explained in the previous chapter, most importantly in lambda abstraction
(the relationship with monads is further discussed in Section~\ref{sec:flat-related}).

@@

The reason for this asymmetry follows from the fact that the context appears in a \emph{negative
position} in the semantic model (Section~\ref{sec:flat-semantics}). It means that we need to be
careful about using the words \emph{split} and \emph{merge}, because they can be read as meaning
exactly the opposite things. To disambiguate, we always use the term \emph{context requirements} 
when using the syntactic view, especially in the rest of Chapter~\ref{chf:flat}, and 
\emph{context capabilities} or just \emph{available context} when using the semantic view, 
especially in Chapter~\ref{ch:transl}.

@@

Furthermore, the semantics of context-aware languages in Chapter~\ref{sec:applications} and
also in Chapter~\ref{sec:transl} is defined over \emph{typing derivation} and so the same program
could have a different meaning, depending on the typing derivation chosen. In this section, we
specify how to choose \emph{unique} typing derivation in each of the coeffect systems we consider.

@@

\paragraph{Properties.}
If a program written in a coeffect language with implicit parameters is well-typed according to the 
type system presented in Figure~\ref{fig:flat-resolve-impl}, then the type system gives a unique
derivation. We use this unique typing derivation to give the semantics of coeffect language
with implicit parameters in Chapter~\ref{ch:transl} and we also implement this algorithm as
discussed in Chatper~\ref{ch:impl}.

@@

\paragraph{Properties.}
If a program written in a coeffect language for liveness or dataflow is well-typed according to
the type system presented in Figure~\ref{fig:flat-types} with the (\emph{abs}) rule replaced by
(\emph{idabs}), then the type system gives a unique derivation. As for implicit parameters, this
defines the semantics of coeffect program (Chapter~\ref{ch:transl}) and it is used in the
implementation (Chatper~\ref{ch:impl}).

@@

The focus of this chapter is on the general coeffect system and so we do not discuss the operational
semantics of the specific notions of context. However, some work in that area has been done
by Brunel et al. \cite{coeffects-quantitative}. We discuss concrete semantics of implicit 
parameters and dataflow in Chapter~\ref{ch:transl}.

>>>>

The development in this chapter can be seen as a counterpart to the well-known development of 
\emph{effect systems} \cite{effects-gifford}. The Chapter~\ref{ch:semantics} then links \emph{coeffects} 
with \emph{comonads} in the same way in which effect systems can be linked with monads
\cite{monad-notions}. The syntax and type system of the flat coeffect calculus follows 
a similar style as effect systems \cite{effects-polymorphic,effects-talpin-et-al}, but differs
in the structure, as explained in the previous chapter, most importantly in lambda abstraction
(the relationship with monads is further discussed in Section~\ref{sec:flat-related}).

@@

The reason for this asymmetry follows from the fact that the context appears in a \emph{negative
position} in the semantic model (Section~\ref{sec:flat-semantics}). It means that we need to be
careful about using the words \emph{split} and \emph{merge}, because they can be read as meaning
exactly the opposite things. To disambiguate, we always use the term \emph{context requirements} 
when using the syntactic view, especially in the rest of Chapter~\ref{chf:flat}, and 
\emph{context capabilities} or just \emph{available context} when using the semantic view, 
especially in Chapter~\ref{ch:semantics}.

@@

Furthermore, the semantics of context-aware languages in Chapter~\ref{sec:applications} and
also in Chapter~\ref{ch:semantics} is defined over \emph{typing derivation} and so the same program
could have a different meaning, depending on the typing derivation chosen. In this section, we
specify how to choose \emph{unique} typing derivation in each of the coeffect systems we consider.

@@

\paragraph{Properties.}
If a program written in a coeffect language with implicit parameters is well-typed according to the 
type system presented in Figure~\ref{fig:flat-resolve-impl}, then the type system gives a unique
derivation. We use this unique typing derivation to give the semantics of coeffect language
with implicit parameters in Chapter~\ref{ch:semantics} and we also implement this algorithm as
discussed in Chatper~\ref{ch:impl}.

@@

\paragraph{Properties.}
If a program written in a coeffect language for liveness or dataflow is well-typed according to
the type system presented in Figure~\ref{fig:flat-types} with the (\emph{abs}) rule replaced by
(\emph{idabs}), then the type system gives a unique derivation. As for implicit parameters, this
defines the semantics of coeffect program (Chapter~\ref{ch:semantics}) and it is used in the
implementation (Chatper~\ref{ch:impl}).

@@

The focus of this chapter is on the general coeffect system and so we do not discuss the operational
semantics of the specific notions of context. However, some work in that area has been done
by Brunel et al. \cite{coeffects-quantitative}. We discuss concrete semantics of implicit 
parameters and dataflow in Chapter~\ref{ch:semantics}.
